Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2014-10-24 Report/wsrm_formulation.tex
updated final with kkt conditions
Submitted version

In this approach, we present an algorithm to solve \eqref{eqn-3} to obtain the transmit precoders in a centralized manner by using the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \tilde{N}_0 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \text{and } \eqref{eqn-4.3}, \IEEEyessubnumber
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \tilde{N}_0 + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \text{and } \eqref{eqn-4.3}, \IEEEyessubnumber
\end{IEEEeqnarray}
}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} is the vector which needs to be identified for the optimal allocation and \me{\tilde{N}_0 = N_0\|\mvec{w}{l,k,n}\|^2} be the effective noise variance. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step leads to the same solution without loss of optimality, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by variable \me{\beta_{l,k,n}}. Similar to the \ac{WSRM} problem in \cite{tran2012fast}, the problem can be shown to be NP-hard even for the single antenna case. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. Recall that the original coordinate descent method assumes that the optimization variables belong to disjoint sets (Cartesian product of sets, to be precise) \cite{xu2013block}.

@@

The proposed algorithm is referred to as \acl{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint, and it is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. For a fixed receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{SCA} iteration is carried out until convergence or for the predefined iterations, say, \me{J_{\max}} for the optimal transmit precoders \me{\mvec{m}{l,k,n}}. Next, the receive beamformers are updated based on either \eqref{opt-rx} or \eqref{eqn-10} using the fixed transmit precoders \me{\mvec{m}{l,k,n}}. This procedure is carried out until convergence of the queue deviation or for fixed number of iterations by \me{I_{\max}} as outlined in Algorithm \ref{algo-1}.
{\allowdisplaybreaks
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and transmit precoders \me{\tilde{\mbf{M}}_{k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{W}{k,n}} and \me{\tilde{\mbf{u}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-8} using \me{\tilde{\mbf{M}}_{k,n}}\;
 \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
 initialize \me{j = 0}\;
 \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
 update the constraint set \eqref{eqn-8} with \me{\tilde{\mbf{u}}_{l,k,n}} and \me{\mvec{m}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
 $j = j + 1$\;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
 $i = i + 1$\;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}
}
\subsubsection*{Convergence}

In order to prove the convergence of the proposed iterative algorithm, following conditions are to be satisfied \cite{scutari}
\begin{itemize}
	\item convergence of the \ac{SCA} subproblem	
	\item uniqueness of the transmit and the receive beamformers
	\item monotonic convergence of the objective function
\end{itemize}
In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, which is majorized by the quadratic-over-linear function in \eqref{eqn-6.2} from below around a fixed point \me{\tilde{\mbf{u}}^{(i)}_{l,k,n}}. Since the \ac{SCA} method is adopted in the proposed algorithm, the constraint approximation satisfies the following conditions as in \cite{marks1978technical}
\begin{subeqnarray} \label{sca-req}
	f(\tilde{\mbf{u}}_{l,k,n}) &\leq& \bar{f}(\tilde{\mbf{u}}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	\nabla f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \nabla \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}),
\end{subeqnarray}
where \me{\bar{f}(\mbf{x},\mbf{x}^{(i)})} is the approximate function of \me{f(\mbf{x})} around the point \me{\mbf{x}^{\ast(i)}}. The stationary point of the relaxed convex problem satisfies the \ac{KKT} conditions of the original nonconvex problem, which can be obtained by using conditions in \eqref{sca-req}. It can be seen that the  \ac{SCA} relaxed formulation converges to a local stationary point at each iteration.

>>>>

In this approach, we present an algorithm to solve \eqref{eqn-3} to obtain the transmit precoders in a centralized manner by using the idea of alternating optimization and successive convex approximation. Using \eqref{eq:SINR}, we can reformulate the problem defined in \eqref{eqn-3} as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\ \beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} & \quad & \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \\
\text{subject to}& \quad &\gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}  |^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \label{eqn-6.2} \\
  & \quad & \beta_{l,k,n} \geq  \tilde{N}_0 + \hspace{-0.75em} \sum_{(j,i) \neq (l,k)} \hspace{-0.75em} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \label{eqn-6.3} \\
  & \quad & \text{and } \eqref{eqn-4.3}, \IEEEyessubnumber
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CCl}\label{eqn-6}
	\underset{\substack{\gamma_{l,k,n},\mvec{m}{l,k,n},\\\beta_{l,k,n},\mvec{w}{l,k,n}}}{\text{minimize}} &\quad& \|  \tilde{\mbf{v}}  \|_q \label{eqn-obj} \IEEEyessubnumber \vspace{-0.25cm}\\
	\text{subject to} &\quad& \gamma_{l,k,n} \leq \frac{ | \mvec{w}{l,k,n}^\herm \mvec{H}{l,k,n} \mvec{m}{l,k,n}|^2}{\beta_{l,k,n}} \triangleq f(\tilde{\mbf{u}}_{l,k,n}) \IEEEyessubnumber \eqspace \label{eqn-6.2} \\
	&\quad& \beta_{l,k,n} \geq \tilde{N}_0 + \sum_{\mathclap{(j,i) \neq (l,k)}} |\mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} |^2 \IEEEyessubnumber \eqspace \label{eqn-6.3} \\
	&\quad& \text{and } \eqref{eqn-4.3}, \IEEEyessubnumber
\end{IEEEeqnarray}
}
where \me{\tilde{\mbf{u}}_{l,k,n} \triangleq \{\mvec{w}{l,k,n}^\herm, \mvec{H}{b_k,k,n}, \mvec{m}{l,k,n},\beta_{l,k,n}\}} is the vector which needs to be identified for the optimal allocation and \me{\tilde{N}_0 = N_0\|\mvec{w}{l,k,n}\|^2} be the effective noise variance. In this formulation, we relaxed the equality constraint in \eqref{eq:SINR} by the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3}. However, this step leads to the same solution without loss of optimality, since the inequalities in \eqref{eqn-6.2} and \eqref{eqn-6.3} are active for an optimal solution, following the same arguments as those in \cite{tran2012fast}. Intuitively, \eqref{eqn-6.2} denotes the \ac{SINR} constraint for \me{\gamma_{l,k,n}}, and \eqref{eqn-6.3} gives an upper bound for the total interference seen by user \me{k \in \mathcal{U}_b}, denoted by the optimization variable \me{\beta_{l,k,n}}.

Similar to the \ac{WSRM} problem in \cite{tran2012fast}, the proposed formulation in \eqref{eqn-6} can be shown to be NP-hard even for the single antenna case. The reformulation in \eqref{eqn-6} allows a tractable solution as presented below. First, we note that the constraints \eqref{eqn-4.3} are convex with involved variables. Thus, we only need to deal with \eqref{eqn-6.2} and \eqref{eqn-6.3}. Towards this end, we resort to the traditional coordinate descent technique by fixing the linear receivers, and finding the optimal transmit beamformers. Recall that the coordinate descent method assumes that the optimization variables belong to disjoint sets and the problem must be convex for a given set of variables while keeping other variables fixed as discussed in \cite{xu2013block}.

@@

The proposed algorithm is referred to as \acl{QM} \ac{JSFRA} scheme with a per \ac{BS} power constraint, and it is outlined in Algorithm \ref{algo-1}. The iterative procedure repeats until the improvement on the objective is less than a predetermined tolerance parameter or the maximum number of iterations is reached. Instead of initializing \me{\tilde{\mbf{u}}_{l,k,n}} arbitrarily to a feasible point, transmit precoders can also be initialized with any feasible point \me{\tilde{\mbf{m}}_{l,k,n}}, which is then used to find \me{\tilde{\mbf{u}}_{l,k,n}} in an efficient manner as briefed in Algorithm \ref{algo-1}. For a fixed receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{SCA} iteration is carried out until convergence or for the predefined iterations, say, \me{J_{\max}} for the optimal transmit precoders \me{\mvec{m}{l,k,n}}. Next, the receive beamformers are updated based on either \eqref{opt-rx} or \eqref{eqn-10} using the fixed transmit precoders \me{\mvec{m}{l,k,n}}. This procedure is carried out until convergence of the queue deviation or for fixed number of iterations by \me{I_{\max}} as outlined in Algorithm \ref{algo-1}.
{\allowdisplaybreaks
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=0} and transmit precoders \me{\tilde{\mbf{M}}_{k,n}} randomly satisfying the total power constraint \eqref{eqn-4.3}}
 update \me{\mvec{W}{k,n}} and \me{\tilde{\mbf{u}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-8} using \me{\tilde{\mbf{M}}_{k,n}}\;
 \Repeat{Queue convergence or \me{i \geq I_{\max}}}{
 initialize \me{j = 0}\;
 \Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
 solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
 update the constraint set \eqref{eqn-8} with \me{\tilde{\mbf{u}}_{l,k,n}} and \me{\mvec{m}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
 $j = j + 1$\;
 }
 update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
 $i = i + 1$\;
 }
 \caption{Algorithm of \acs{JSFRA} scheme}
 \label{algo-1}
\end{algorithm}
}

\subsubsection*{Convergence}
In order to prove the convergence of the proposed iterative algorithm, following conditions are to be satisfied \cite{scutari}
\begin{itemize}
	\item convergence of the \ac{SCA} subproblem	
	\item uniqueness of the transmit and the receive beamformers
	\item monotonic convergence of the objective function
\end{itemize}
In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, which is majorized by the quadratic-over-linear function in \eqref{eqn-6.2} from below around a fixed point \me{\tilde{\mbf{u}}^{(i)}_{l,k,n}}. Since the \ac{SCA} method is adopted in the proposed algorithm, the constraint approximation satisfies the following conditions as in \cite{marks1978technical}
\begin{subeqnarray} \label{sca-req}
	f(\tilde{\mbf{u}}_{l,k,n}) &\leq& \bar{f}(\tilde{\mbf{u}}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	\nabla f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \nabla \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}),
\end{subeqnarray}
where \me{\bar{f}(\mbf{x},\mbf{x}^{(i)})} is the approximate function of \me{f(\mbf{x})} around the point \me{\mbf{x}^{\ast(i)}}. The stationary point of the relaxed convex problem satisfies the \ac{KKT} conditions of the original nonconvex problem, which can be obtained by using conditions in \eqref{sca-req}. It can be seen that the  \ac{SCA} relaxed formulation converges to a local stationary point at each iteration.
